---
title: "Analyse TopoSUB simulation"
author: "Johannes Brenner, EURAC research"
date: "7. August 2015"
output: html_document
runtime: shiny
---

This R Markdown document is made interactive using Shiny. Unlike the more traditional workflow of creating static reports, you can now create documents that allow your readers to change the assumptions underlying your analysis and see the results immediately. 

To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

---

```{r, include=FALSE}

# load required libs

if(!require(DataBaseAlpEnvEURAC)) {
  
  if(!require(devtools)) {
    install.packages("devtools")
    library(devtools)
  }
  
  install.packages("DataBaseAlpEnvEURAC")
  library(DataBaseAlpEnvEURAC)
}

if(!require(raster)) {
  install.packages("raster")
  library(raster)
}

if(!require(data.table)) {
  install.packages("data.table")
  library(data.table)
}

if(!require(dplyr)) {
  install.packages("dplyr")
  library(dplyr)
}

if(!require(tidyr)) {
  install.packages("tidyr")
  library(tidyr)
}

if(!require(leaflet)) {
  install.packages("leaflet")
  library(leaflet)
}

if(!require(zoo)) {
  install.packages("zoo")
  library(zoo)
}

if(!require(dygraphs)) {
  install.packages("dygraphs")
  library(dygraphs)
}

if(!require(rgdal)) {
  install.packages("rgdal")
  library(rgdal)
}

if(!require(chron)) {
  install.packages("chron")
  library(chron)
}

if(!require(d3heatmap)) {
  install.packages("d3heatmap")
  library(d3heatmap)
}

```

```{r, include=FALSE}

# load data
#simpath <- "/run/user/1000/gvfs/smb-share:server=sdcalp01.eurac.edu,share=data2//Simulations/Simulation_GEOtop_1_225_ZH/Vinschgau/SimTraining/BrJ/Mazia/toposub/sim/1d/1d_001/000004"

simpath <- "/home/jbr/Schreibtisch/topotestdata"

# meta info Mazia stations
data("station_meta")
station_meta <- station_meta[c(2,3,16,18:21),]
xy <- project(xy=cbind(station_meta$x,station_meta$y), proj = "+proj=utm +zone=32 ellps=WGS84", inv = T)
station_meta$long <- xy[,1]
station_meta$lat <- xy[,2]

# toposub setup
setup <- read.csv(file.path(simpath, "setup.txt"), header = FALSE)
Nclust <- setup[setup$V2=="Nclust",3]

locations <- read.csv(file.path(simpath, "locations.txt"), header = FALSE)

# load simulation data - points
load(file.path(simpath, "data.RData"))

dates <- sort(format(as.Date(substr(unique(data_00004$Date12_DDMMYYYYhhmm_),1,10), "%d/%m/%Y"), "%Y-%m-%d"))

# read landform to rst
landform <- raster(x = file.path(simpath, paste("landform_", Nclust, ".asc", sep="")))

landfval <- getValues(landform)
landform[] <- ifelse(landfval==landfval[1], NA, landfval)

elevation <- raster(x = file.path(simpath, "INmaps10", "dem.asc"))

landcover <- raster(x = file.path(simpath, "INmaps10", "landcover.asc"))
#soil <- raster(x = file.path(simpath, "INmaps10", "soil.asc"))

# listpoints table
listpoints <- read.csv(file.path(simpath, "listpoints.txt"))
toDel <- which.max(listpoints$members)
listpoints <- listpoints[listpoints$id!=toDel,]

data_00004 <- data_00004[IDpoint!=toDel,]

```

### Choices



```{r, echo=FALSE}

inputPanel(
  # choose variable
  selectInput(inputId = "Variable", label = "Variable to investigate:",
              choices = names(data_00004)[-c(1:4)], selected = names(data_00004)[-c(1:4)][1]),
  
  # choose point ID
  selectInput(inputId = "IDpoint", label = "ID of point simulation:",
              choices = unique(data_00004$IDpoint), selected = 1),
  
  # choose start day / end day
  dateInput(inputId = "start", label = "Select Start Date", value = "2070-01-01", min = dates[1], max = tail(dates,1),
            startview = "year"),
  dateInput(inputId = "end", label = "Select End Date", value = "2099-12-31", min = dates[1], max = tail(dates,1),
            startview = "year"),
  
  # choose aggregation function 
  selectInput(inputId = "Aggregation", label = "Function to aggregate:",
              choices = c("mean", "sum", "max", "min", "sd"), selected = "mean"),
  
  # smoothing of map?
  radioButtons(inputId = "Interpolate", label = "Smooth Map", choices = c("YES","NO"), selected = "YES", 
               inline = FALSE),
  
  submitButton(text = "Apply Changes", icon("refresh"))
  )
  
  # get data
#   output$date <- renderUI({
#     seq(as.Date(input$start, "%Y-%m-%d"), as.Date(input$end, "%Y-%m-%d"), 1)
#   })
  
#   output$map <- renderUI({
#     varnr <- which(names(data_00004) ==  input$Variable)
#     num_start <- as.numeric(as.Date(input$start, "%Y-%m-%d")); num_end <- as.numeric(as.Date(input$end, "%Y-%m-%d"))
#     
#     data_var <- data_00004[,c(1,4,varnr),with=F]
#     data_var[,Date12_DDMMYYYYhhmm_ := as.numeric(as.Date(substr(Date12_DDMMYYYYhhmm_,1,10), "%d/%m/%Y"))]
#     
#     data_ <- 
#       data_var %>%
#         setnames(old = names(data_var), new = c("Date","IDpoint","VAR")) %>%
#         filter(Date >= num_start & Date <= num_end) %>%
#         group_by(IDpoint) %>%
#         summarise_each(funs_(input$Aggregation))
#     
#     subs_df <- as.data.frame(data_[,c(1,3), with=F])
#     rst <- subs(landform,subs_df)
#     
#     if(input$Aggregation=="YES") rst <- focal(rst, w=matrix(1, 7, 7), mean)
#     
#     rst
#   })
#   
#   output$ts <- renderUI({
#     varnr <- which(names(data_00004) ==  input$Variable)
#     date <- as.Date(substr(data_00004$Date12_DDMMYYYYhhmm_[data_00004$IDpoint==input$IDpoint],1,10), "%d/%m/%Y")  
#     datazoo <- zoo(data_00004[data_00004$IDpoint==input$IDpoint,input$Variable, with=F], date)
#     datazoo_win <- window(datazoo, start = as.Date(input$start,"%Y-%m-%d"), end = as.Date(input$end,"%Y-%m-%d"))
#     datazoo_win
#   )}

```

---

```{r, echo=FALSE}

renderDataTable({
  as.data.table(round(listpoints,2))
}, options = list(pageLength=5, lengthMenu=c(5, 10, 15, 20, 50, 100))
)

```

---

### Mapping anomaly

Anomaly of variable for a specific climate period (see choices above) compared to baseline (1970-2000). 

```{r, echo=FALSE}

# Mapping

renderLeaflet({
    varnr <- which(names(data_00004) ==  input$Variable)
    num_start <- as.numeric(as.Date(input$start, "%Y-%m-%d")); num_end <- as.numeric(as.Date(input$end, "%Y-%m-%d"))

    num_start_baseline <- as.numeric(as.Date("1970-01-01", "%Y-%m-%d"))
    num_end_baseline <- as.numeric(as.Date("1999-12-31", "%Y-%m-%d"))
    
    data_var <- data_00004[,c(1,4,varnr),with=F]
    data_var[,Date12_DDMMYYYYhhmm_ := as.numeric(as.Date(substr(Date12_DDMMYYYYhhmm_,1,10), "%d/%m/%Y"))]
    
    data_ <- 
      data_var %>%
        setnames(old = names(data_var), new = c("Date","IDpoint","VAR")) %>%
        filter(Date >= num_start & Date <= num_end) %>%
        group_by(IDpoint) %>%
        summarise_each(funs_(input$Aggregation))
    
    data_base <- 
      data_var %>%
        setnames(old = names(data_var), new = c("Date","IDpoint","VAR")) %>%
        filter(Date >= num_start_baseline & Date <= num_end_baseline) %>%
        group_by(IDpoint) %>%
        summarise_each(funs_(input$Aggregation))
    
    data_[, dif := data_$VAR - data_base$VAR]
    
    subs_df <- as.data.frame(data_[,c(1,4), with=F])
    rst <- subs(landform,subs_df)
    
    if(input$Interpolate=="YES") rst <- focal(rst, w=matrix(1, 7, 7), mean)
  
  pal1 <- colorNumeric(c("#FFFFCC", "#41B6C4", "#0C2C84"), values(rst), na.color = "transparent")
  
  crs(rst) <- "+proj=utm +zone=32 ellps=WGS84"
  
  lndf <- landform
  lndf[] <- ifelse(landfval==input$IDpoint, values(elevation), NA)
  crs(lndf) <- "+proj=utm +zone=32 ellps=WGS84"
  
  pal2 <- colorNumeric(c("#f03b20", "#feb24c", "#ffeda0"), values(lndf), na.color = "transparent")

  crs(landcover) <- "+proj=utm +zone=32 ellps=WGS84"
  #crs(soil) <- "+proj=utm +zone=32 ellps=WGS84"
  
  leaflet() %>%
    addProviderTiles("Acetate.terrain", group="Terrain") %>%
    addTiles(options = providerTileOptions(opacity = 0.3), group = "OSM") %>%  # Add default OpenStreetMap map tiles
    #addRasterImage(soil, opacity = 0.5, group = "Soil") %>% # Add aggregated raster
    addRasterImage(landcover, opacity = 0.5, group = "Landcover") %>% # Add aggregated raster
    addRasterImage(rst, colors = pal1, opacity = 0.75, group = "Raster") %>% # Add aggregated raster
    addRasterImage(lndf, colors = pal2, opacity = 0.4, group = "Landform") %>% # Add aggregated raster
    addLegend(pal = pal1, values = values(rst), title = input$Variable) %>%
    #addLegend(pal = pal2, values = values(lndf), title = "m a.s.l.") %>%
    addCircleMarkers(lng=station_meta$long, lat=station_meta$lat, popup=paste(station_meta$name, ", ",  station_meta$h, "m a.s.l"), fill = F, group = "Stations") %>%
    addLayersControl(position = "topleft" , overlayGroups = c("Terrain", "OSM", "Landcover", "Raster", "Landform", "Stations"),
    options = layersControlOptions(collapsed = FALSE))
  
})

```

---

### Time series

Time series on daily scale and yearly anomalies to baseline mean/sum.

```{r, echo=FALSE}

# TimeSeries pointID

renderDygraph({
  
    varnr <- which(names(data_00004) ==  input$Variable)
    date <- as.Date(substr(data_00004$Date12_DDMMYYYYhhmm_[data_00004$IDpoint==input$IDpoint],1,10), "%d/%m/%Y")  
    datazoo <- zoo(data_00004[data_00004$IDpoint==input$IDpoint,input$Variable, with=F], date)
 #   datazoo_win <- window(datazoo, start = as.Date(input$start,"%Y-%m-%d"), end = as.Date(input$end,"%Y-%m-%d"))
  
  dygraph(datazoo) %>%
    dyRangeSelector() %>%
    dyRoller()
})

renderPlot({
  
    varnr <- which(names(data_00004) ==  input$Variable)
    date <- as.Date(substr(data_00004$Date12_DDMMYYYYhhmm_[data_00004$IDpoint==input$IDpoint],1,10), "%d/%m/%Y")  
    datazoo <- zoo(data_00004[data_00004$IDpoint==input$IDpoint,input$Variable, with=F], date)
    
    #yearly data
    datazoo_yr <- aggregate(datazoo, years(time(datazoo)), FUN = input$Aggregation)
    datazoo_yr <- datazoo_yr[-dim(datazoo_yr)[[1]],]
    
    datazoo_yr_dev <- datazoo_yr - mean(window(datazoo_yr, start = 1970, end=2000))
    
      # linear filter
    op <- par(mfrow=c(1,1), las=1)
    k = c(.5,1,1,1,.5)            # k is the vector of weights
    k = k/sum(k)
    fTy_dev <- stats::filter(ts(datazoo_yr_dev), sides=2, k)
    plot(datazoo_yr_dev, type="h", ylab=paste("deviation", input$Variable), xlab="YEAR")
    points(datazoo_yr_dev, pch=16, cex=.75)
    lines(fTy_dev, col="red")
    lines(lowess(datazoo_yr_dev), col="blue", lty="dashed", lwd=3)
    abline(h=c(-1,0,1,2))
    par(op)

})

```

---

### Heatmap

Yearly anomalies to baseline mean/sum.

```{r, echo=FALSE}

renderD3heatmap({
  # prepare listpoints
  names(listpoints)[2] <- "IDpoint"
  
  # baseline start: 1980-01-01
  winstart <- as.numeric(as.Date("1980-01-01")); winend <- as.numeric(as.Date("2010-12-31"))
  
  date <- as.Date(substr(data_00004$Date12_DDMMYYYYhhmm_,1,10), "%d/%m/%Y")
  datenum <- as.numeric(date)
  year    <- years(date)
  months  <-  as.numeric(format(date, "%m"))
  data_00004[,"datenum"] <- datenum
  data_00004[,"year"]    <- year
  data_00004[,"month"]   <- months
  
  if (grepl("SWC", input$Variable))   {
    data_base <- data_00004[datenum>winstart & datenum<winend & month > 3 & month < 11, c("IDpoint",input$Variable), with=F]
  } else {
    data_base <- data_00004[datenum>winstart & datenum<winend, c("IDpoint",input$Variable), with=F]
  }
  
  setnames(data_base, names(data_base), c("IDpoint", "VAR"))
  
data_base_summ <-   
  data_base %>%
    group_by(IDpoint) %>%
    summarise(meanVAR = mean(VAR)) %>%
    merge(listpoints, by="IDpoint")

  data <- data_00004[, c("year","IDpoint",input$Variable), with=F]
  setnames(data, names(data), c("year", "IDpoint", "VAR"))

data_year_mean <- 
  data %>% 
    group_by(year,IDpoint) %>%
    summarise(meanVAR=mean(VAR)) %>%
    spread(year,meanVAR)
 
mat <- as.matrix(data_year_mean[, names(data_year_mean)[-1], with=F])
vec <- c(data_base_summ[,"meanVAR",  with=F])$meanVAR
var_dev  <- (mat - vec)
var_perc <- var_dev / vec *100

cellnote <- matrix(paste(as.character(round(var_perc,1)), "%; ", as.character(round(var_dev,3)), " UNITs",sep=""), nrow = nrow(var_perc) ,ncol = ncol(var_perc))

#d3heatmap(var_dev[,-dim(var_dev)[[2]]], Rowv=T, Colv=F, scale = "none", cellnote = cellnote[,-dim(var_dev)[[2]]], labRow = paste(listpoints$landcover, "-", round(listpoints$slp,0), "-", round(listpoints$dem,0), sep=" "), dendrogram = "row")

d3heatmap(var_perc[,-dim(var_perc)[[2]]], Rowv=T, Colv=F, scale = "none", cellnote = cellnote[,-dim(var_dev)[[2]]], labRow = paste(listpoints$landcover, "-", round(listpoints$slp,0), "-", round(listpoints$dem,0), sep=" "), dendrogram = "row")


})

```
