---
title: 'Lect 2: Postprocess a TopoSUB simulation'
author: "Johannes Brenner"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Lect 2 Postprocess}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Before starting make sure that the [TopoSUB R-package](https://github.com/JBrenn/TopoSUB) is installed and can be loaded:

```{r libraries, eval=FALSE}

if(!require(TopoSUB))
{
  if(!require("devtools"))
  {
    install.packages(devtools)
    require("devtools")
  }
  install_github("TopoSUB", "JBrenn")
  require("TopoSUB")
}

```

## Main functions

The main functions for TopoSUB postprocessing are reading and re-maping the cluster centroids of the output variables of a TopoSUB simulation. Look at their descripton!

```{r help, eval=FALSE}

?TopoSUB_read
?TopoSUB_remap

```

1. **TopoSUB_read** is reading the TopoSUB simulation output data in a data.table, it is utilising the R-function _data.table::fread_ for faster and more convenient reading of text files compared to _read.table_, for simulations run in parallel mode parallel reading of data is implemented. Some postprocessing is done for specific variables:

* calculation of total precipitation from rain and snow components,
* calculation of evapotranspiration from transpiration and evaporation,
* calculation of latent and sensible heat over the canopy (see function _TopoSUB_EfluxOcanopy_),
* identification of snow cover days defined by the threshold argument _SnowCoverThres_,
* calculation of total soil water content by adding liquid and ice soil water content for each soil layer.

2. **TopoSUB_remap** is taking over the data.table object (argument _data_) and is re-mapping variable vise (argument _variable_) according to the TopoSUB output map _landform .asc_. Variable names to use can be obtained by sourcing _colnames_ from the TopoSUB package:

```{r colnames, eval=FALSE}

data("colnames")
View(colnames)

```

The further arguments are a bit more complex and shortly described below:

* **periods** is a list defining the periods which are compared, at least two time periods Have to be set, preferable climatic time frames (30 years). This is a named list, which first entry **baseline** is substracted from the other periods. Every period is aggregated over the whole time frame, the aggregation is defined in the argument

* **periods_aggr**, which is a named list of three entries: **aggr** defines the time aggregation, e.g. seasons (_season_), vegetation period (_veg_period_), by default an aggregation over the whole year is always performed. **fun** defines the aggregation function, which is either the _mean_ or the _sum_ of the variable  over the aggragation period. **diff** defines if the _absolute_ or _percentage_ difference between the periods is computed.

* **sequence** is meant to produce map animations over a specific time period. It's not implemented yet.

* **postprocess** is defining a further postprocess step for specific variables, e.g. number of drought days calculated from soil water content, see function _TopoSUB_POSTcritSWC_fc_wp_. The character after POST is the keyword for the _method_, the first entry of the named vector, the following entries are extra arguments for the postprocess function.

For clarification see the definitions of the arguments below:

```{r periods, eval=FALSE}

periods = list(baseline = c(1980,2010), per1 = c(2020,2050), 
                            per2 = c(2045,2075), per3 = c(2070,2100))

periods_aggr = list(aggr = c("season", "veg_period"), fun = "mean", 
                    diff = c("absolute","percentage"))

postprocess = c(method="critSWC_fc_wp", dry_thres="30")

```


## Example script

```{r example, eval=FALSE}
#==============================================================================
# run TopoSUB postprocessing
#==============================================================================

# set working directory to TopoSUB simulation folder

# can also be run with working directory in root TopoSUB folder
# if location configuration didn't change

# load parallel library
  library(parallel)
  
# setup file
  setupfile = "setup.txt"
  
# read location file
  locationfile = "locations.txt"
  locations <- read.csv(locationfile, header = F, colClasses="character")
  
  # get parameter
  apply(X = locations[,c(2,3)], MARGIN = 1,
        FUN = function(x) assign(x = x[1], value = x[2], envir = .GlobalEnv) )

# get sim path from location file
  simpath <- file.path(root, sim, "1d", paste("1d_",exper1,sep=""), 
                       formatC(as.integer(es),0,6,flag = "0"))
  
# define selection  argument  
  select <- list(PointOutputFileWriteEnd=c("Date12[DDMMYYYYhhmm]", "IDpoint", "Tair[C]", "Prain_over_canopy[mm]", "Psnow_over_canopy[mm]", "snow_water_equivalent[mm]", "Evap_surface[mm]", "Trasp_canopy[mm]", "Hv[W/m2]", "LEv[W/m2]", "Hg_unveg[W/m2]", "LEg_unveg[W/m2]", "Hg_veg[W/m2]", "LEg_veg[W/m2]", "Canopy_fraction[-]"), 
                 SoilLiqContentProfileFileWriteEnd=c("Date12[DDMMYYYYhhmm]", "IDpoint", "20.000000", "50.000000", "200.000000",  "500.000000"),
                 SoilIceContentProfileFileWriteEnd=c("Date12[DDMMYYYYhhmm]", "IDpoint", "20.000000", "50.000000", "200.000000", "500.000000"))
  
# read output
  data <- TopoSUB_read(wpath=simpath,
                       keys= c("PointOutputFileWriteEnd","SoilLiqContentProfileFileWriteEnd", 
                               "SoilIceContentProfileFileWriteEnd"),
                       select = select)
 
# re-mapping for long term climate impact simulations
  
  # list defining variable, fun and postprocess
  
  var_in <- list(c("Evaptranspiration_mm_","sum",NULL), 
                 c("Ptotal_mm_","sum",NULL),
                 c("snow_water_equivalent_mm_","sum",NULL), 
                 c("SnowCover_days_","sum",NULL),
                 c("SWC_total_200","mean",NULL), 
                 c("SWC_ice_200","sum",method=NULL),
                 c("SWC_total_200","sum",method="critSWC_fc_wp",dry_thres="30"))
  
  # parallelisation - looping through var_in with mclapply
  # applying TopoSUB_remap function
  
  mclapply(var_in, function(x) {
    TopoSUB_remap(data = data, variable = x[1], wpath = simpath,
                  periods_aggr = list(aggr=c("season", "veg_period"), fun=x[2], 
                                      diff = c("absolute","percentage")),
                  sequence=NULL, postprocess=x[3:length(x)])
  }, data=data, simpath=simpath)

```


## Output in the TopoSUB simulation folder

* _OUTcsv_ containing monthly and yearly timeseries (.csv) for each cluster according to the aggergation and postprocess you applied. 

* _OUTperiods_ containing maps (GEOtiff) for all periods and difference maps according to the aggregation settings applied.

### Requirements

* current installation of R (>3.0.0)
* [gdal](http://www.gdal.org/) and [proj4](https://github.com/OSGeo/proj.4) installed
