---
title: 'Lect 3: How to Visualize TopoSUB output'
author: "Johannes Brenner"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Lect 3 Visualisation}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Get GitHub Repository TopoSUB

Before starting download and unzip OR clone the [GitHub repository TopoSUB](https://github.com/JBrenn/TopoSUB):

```{r, engine='bash', eval=FALSE}
git clone https://github.com/JBrenn/TopoSUB
```

Copy the files within the shiny folder (ui.R, server.R, runApp.R) in your TopoSUB simulation folder.

## Dependencies

Before running the app make sure you have installed the following libraries

1. [data.table](https://cran.r-project.org/web/packages/data.table/index.html) an extension of data.frame - a table plug-in for jQuery,
2. [tidyr](https://blog.rstudio.org/2014/07/22/introducing-tidyr/) makes it easy to “tidy” your data. Tidy data is data that’s easy to work with: it’s easy to munge (with dplyr), visualise (with ggplot2), ...
3. [dplyr](https://blog.rstudio.org/2014/01/17/introducing-dplyr/) providing a set of tools for efficiently manipulating datasets in R,
4. [raster](https://cran.r-project.org/web/packages/raster/index.html) for Geographic Data Analysis and Modeling,
5. [ggplot2](http://ggplot2.org/) ggplot2 is a plotting system for R, based on the grammar of graphics, which tries to take the good parts of base and lattice graphics and none of the bad parts, 
6. [leaflet](http://leafletjs.com/) is one of the most popular open-source JavaScript libraries for interactive maps,
7. [shiny](http://shiny.rstudio.com/) a web application framework for R.

```{r libraries, eval=FALSE}

install.packages(c("data.table","tidyr","dplyr","raster","ggplot2","leaflet","shiny"))

```

## Running the shiny application 

Close RStudio, and double-click _runApp.R_ (in the TopoSUB simulation folder). This should open the R script in RStudio with your TopoSUB simulation folder as a working directory. The working directory is written above the R console, you can also check with

```{r getwd, eval=FALSE}

getwd()

```

Execute the whole script: 

```{r shinyApp, eval=FALSE}

library(shiny)
library(leaflet)
library(raster)
library(data.table)
library(tidyr)
library(dplyr)
library(ggplot2)

runApp(getwd())

```

A browser window should open and show the shiny app interface. If your data is stored on a server reading map files is slow. You can improve speed by copying the simuation folder (at least the folders _OUTperiods_ and _OUTcsv_, as well as the file _listpoints.txt_) on your local machine.

### Requirements

* current installation of R (>3.0.0) and RStudio Desktop (>0.99)
* TopoSUB simulation run (see Lect 1: How to Run TopoSUB) and postprocessing (see Lect 2: How to Postprocess a TopoSUB sim)
